---
title: "Dashboard"
format: dashboard
editor_options: 
  chunk_output_type: console
execute:
  echo: false
---

```{r}
#| label: load-packages
#| message: false

# Required packages
pacman::p_load(
   tidyverse     # Data manipulation and visualization
  ,dlookr        # Data diagnosis and exploration
  ,kableExtra    # Enhanced 'knitr::kable()' tables
  ,scales        # Visualization scaling tools
  ,gt            # Stylish display tables
  ,gridExtra     # Arrange plots
  ,googleAnalyticsR  # Google Analytics API
)
```

```{r}
#| label: load-data

# Load dataset
file_path <- "data/Different_stores_dataset.csv"
data <- read.csv(file_path)

# Convert date columns to Date format
data$order_date <- as.Date(data$order_date, format = "%m/%d/%Y")
data$ship_date <- as.Date(data$ship_date, format = "%m/%d/%Y")
```

```{r}
#| label: ga-authentication
#| message: false
#| warning: false

# Google Analytics authentication
# Replace with your Measurement ID
GA_MEASUREMENT_ID <- "G-Y2163YKVWQ"

# Authenticate with Google Analytics using service account credentials
tryCatch({
  ga_auth(json_file = "ga-credentials.json")
}, error = function(e) {
  message("Note: GA authentication not available. Using demo data.")
})
```

```{r}
#| label: fetch-ga-data
#| message: false
#| warning: false

# Try to fetch GA data if authentication successful
ga_data <- NULL
is_demo <- FALSE

tryCatch({
  # Fetch data from the past 30 days
  ga_data <- google_analytics_4(
    GA_MEASUREMENT_ID,
    date_range = c(Sys.Date() - 30, Sys.Date()),
    metrics = c("activeUsers", "sessions", "screenPageViews"),
    dimensions = c("country", "deviceCategory", "operatingSystem"),
    limit = 10000
  )
  
  if (nrow(ga_data) == 0) {
    is_demo <- TRUE
    message("No GA data found yet. Using demo data.")
  }
}, error = function(e) {
  message("Note: GA data not available. Using demo data instead.")
  is_demo <<- TRUE
})

# Create demo data if GA data not available
if (is.null(ga_data) || nrow(ga_data) == 0 || is_demo) {
  ga_data <- tibble(
    country = c("United States", "Canada", "United Kingdom", "India", "Germany", "Australia", "France", "Japan", "Brazil", "Mexico"),
    deviceCategory = c("mobile", "desktop", "mobile", "desktop", "tablet", "mobile", "desktop", "mobile", "desktop", "tablet"),
    operatingSystem = c("Android", "Windows", "iOS", "Windows", "iPadOS", "Android", "Windows", "iOS", "macOS", "iPadOS"),
    activeUsers = c(145, 320, 87, 156, 42, 98, 203, 71, 134, 56),
    sessions = c(210, 450, 120, 220, 65, 140, 310, 105, 195, 85),
    screenPageViews = c(890, 1250, 310, 680, 180, 420, 950, 280, 710, 240)
  )
  is_demo <- TRUE
}
```

# Overall Metrics

```{r}
#| label: overall-values
# <!-- #| results: hide -->

# Define value box values
n_orders <- nrow(data)
total_sales <- sum(data$sales, na.rm = TRUE)
total_profit <- sum(data$profit, na.rm = TRUE)
avg_discount <- mean(data$discount, na.rm = TRUE)
profit_margin_rate <- total_profit / total_sales
```

## Overall Metrics {style="text-align: center; margin-bottom: 20px; font-size: 28px; font-weight: bold; color: #333;"}

## Row {height="20%" style="display: flex; justify-content: space-evenly; align-items: center;"}

```{r}
#| content: valuebox
#| title: "Total Orders"
#| title-style: "font-size: 24px; font-weight: bold; text-align: center;"

list(
  color = "#3498db",
  value = format(n_orders, big.mark = ","),
  value_style = "font-size: 30px; font-weight: bold;"
)
```

```{r}
#| content: valuebox
#| title: "Total Sales"
#| title-style: "font-size: 24px; font-weight: bold; text-align: center;"

list(
  color = "#3498db",
  value = scales::dollar(total_sales),
  value_style = "font-size: 30px; font-weight: bold;"
)
```

```{r}
#| content: valuebox
#| title: "Total Profit"
#| title-style: "font-size: 24px; font-weight: bold; text-align: center;"

list(
  color = "#3498db",
  value = scales::dollar(total_profit),
  value_style = "font-size: 30px; font-weight: bold;"
)
```

```{r}
#| content: valuebox
#| title: "Average Discount"
#| title-style: "font-size: 24px; font-weight: bold; text-align: center;"

list(
  color = "#3498db",
  value = scales::percent(avg_discount),
  value_style = "font-size: 30px; font-weight: bold;"
)
```

```{r}
#| content: valuebox
#| title: "Profit Margin Rate"
#| title-style: "font-size: 24px; font-weight: bold; text-align: center;"

list(
  color = "#3498db",
  value = scales::percent(profit_margin_rate),
  value_style = "font-size: 30px; font-weight: bold;"
)
```

## Visualizations

## Row {height="80%"}
### Column {width="33%"} {.tabset}

```{r}
#| title: Sales By Region
#<!-- #| fig-height: 8 -->
#<!-- #| fig-width: 12 -->

sales_region <- data %>% 
  group_by(region) %>% 
  summarize(total_sales = sum(sales, na.rm = TRUE))

ggplot(sales_region, aes(x = reorder(region, total_sales), y = total_sales, fill = region)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_y_continuous(labels = scales::dollar) +
  labs(x = "Region", y = "Total Sales") +
  theme(
        axis.text = element_text(size = 36),
        axis.title = element_text(size = 46))
```

### Column {width="33%"} {.tabset}

```{r}
#| title: Profit By Category
#<!-- #| fig-height: 8 -->
#<!-- #| fig-width: 12 -->

profit_category <- data %>% 
  group_by(category) %>% 
  summarize(avg_profit_margin = mean(profit_margin, na.rm = TRUE))

ggplot(profit_category, aes(x = reorder(category, avg_profit_margin), y = avg_profit_margin, fill = category)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_y_continuous(labels = scales::percent) +
  labs( x = "Category", y = "Average Profit Margin") +
  theme(axis.text = element_text(size = 36),
        axis.title = element_text(size = 46))
```

### Column {width="34%"} {.tabset}

```{r}
#| title: Top 10 Subcategories by Sales
#<!-- #| fig-height: 8 -->
#<!-- #| fig-width: 12 -->

sales_subcategory <- data %>% 
  group_by(subcategory) %>% 
  summarize(total_sales = sum(sales, na.rm = TRUE)) %>% 
  top_n(10, total_sales)

ggplot(sales_subcategory, aes(x = reorder(subcategory, total_sales), y = total_sales, fill = subcategory)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_y_continuous(labels = scales::dollar) +
  labs( x = "Subcategory", y = "Total Sales") +
  theme(axis.text = element_text(size = 36),
        axis.title = element_text(size = 46))
```

# Website Analytics

```{r}
#| label: ga-note
if (is_demo) {
  cat("**Note:** Displaying demo data. Real Google Analytics data will appear after 24-48 hours of website traffic.")
}
```

## Overall Web Metrics

## Row {height="20%" style="display: flex; justify-content: space-evenly; align-items: center;"}

```{r}
#| content: valuebox
#| title: "Active Users (Last 30 days)"
#| title-style: "font-size: 24px; font-weight: bold; text-align: center;"

list(
  color = "#2ecc71",
  value = format(sum(ga_data$activeUsers, na.rm = TRUE), big.mark = ","),
  value_style = "font-size: 30px; font-weight: bold;"
)
```

```{r}
#| content: valuebox
#| title: "Total Sessions"
#| title-style: "font-size: 24px; font-weight: bold; text-align: center;"

list(
  color = "#2ecc71",
  value = format(sum(ga_data$sessions, na.rm = TRUE), big.mark = ","),
  value_style = "font-size: 30px; font-weight: bold;"
)
```

```{r}
#| content: valuebox
#| title: "Total Page Views"
#| title-style: "font-size: 24px; font-weight: bold; text-align: center;"

list(
  color = "#2ecc71",
  value = format(sum(ga_data$screenPageViews, na.rm = TRUE), big.mark = ","),
  value_style = "font-size: 30px; font-weight: bold;"
)
```

```{r}
#| content: valuebox
#| title: "Avg. Pages per Session"
#| title-style: "font-size: 24px; font-weight: bold; text-align: center;"

avg_pages <- if(sum(ga_data$sessions, na.rm = TRUE) > 0) {
  round(sum(ga_data$screenPageViews, na.rm = TRUE) / sum(ga_data$sessions, na.rm = TRUE), 2)
} else {
  0
}

list(
  color = "#2ecc71",
  value = avg_pages,
  value_style = "font-size: 30px; font-weight: bold;"
)
```

## User Demographics

## Row {height="80%"}

### Column {width="50%"} {.tabset}

```{r}
#| title: Users by Country

ga_country <- ga_data %>%
  group_by(country) %>%
  summarize(
    activeUsers = sum(activeUsers, na.rm = TRUE),
    sessions = sum(sessions, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(desc(activeUsers)) %>%
  head(10)

ggplot(ga_country, aes(x = reorder(country, activeUsers), y = activeUsers, fill = country)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Top 10 Countries by Active Users",
    x = "Country",
    y = "Active Users"
  ) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold"),
    title = element_text(size = 16, face = "bold")
  )
```

### Column {width="50%"} {.tabset}

```{r}
#| title: Users by Device Type

ga_device <- ga_data %>%
  group_by(deviceCategory) %>%
  summarize(
    activeUsers = sum(activeUsers, na.rm = TRUE),
    sessions = sum(sessions, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(desc(activeUsers))

ggplot(ga_device, aes(x = reorder(deviceCategory, activeUsers), y = activeUsers, fill = deviceCategory)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Active Users by Device Type",
    x = "Device Type",
    y = "Active Users"
  ) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold"),
    title = element_text(size = 16, face = "bold"),
    plot.background = element_rect(fill = "white")
  )
```

## Row {height="80%"}

### Column {width="100%"}

```{r}
#| title: Operating System Distribution

ga_os <- ga_data %>%
  group_by(operatingSystem) %>%
  summarize(
    activeUsers = sum(activeUsers, na.rm = TRUE),
    sessions = sum(sessions, na.rm = TRUE),
    screenPageViews = sum(screenPageViews, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(desc(activeUsers)) %>%
  head(8)

ggplot(ga_os, aes(x = reorder(operatingSystem, activeUsers), y = activeUsers, fill = operatingSystem)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Active Users by Operating System (Top 8)",
    x = "Operating System",
    y = "Active Users"
  ) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold"),
    title = element_text(size = 16, face = "bold")
  )
```
